Letter: Performance models for \\ high-resolution ice sheet simulations

### 2022-06-22 paper/perfmod.tex
update titles and tables

\title[Performance models of ice sheet simulations]{Letter: Performance models of ice sheet simulations}

@@

\begin{table*}[ht]
\begin{center}
\begin{tabular}{cll}
\emph{name} & \emph{meaning} & \emph{units} \\ \hline
$D$         & representative geometric (shallow ice approximation) diffusivity of an ice sheet & $\text{km}^2 \text{a}^{-1}$ \\
$\delta$    & one velocity solution requires $O(n^{1+\delta})$ work\\
$\eps$      & one geometry-update solution requires $O(n^{1+\eps})$ work \\
$L$         & width of simulation domain & km \\
$m$         & total number of stability-limited time steps in actual simulation \\
$m_c$       & total number of time steps needed to resolve modeled climate interactions ($m_c=qT$) \\
$n$         & degrees of freedom: number of horizontal grid points \\
$q$         & time steps per model year needed to resolve modeled climate interactions & $\text{a}^{-1}$ \\
$T$         & simulation duration in model years & a \\
$\Delta t$  & length of time step in model years & a \\
$U$         & representative horizontal ice velocity & $\text{km}\,\text{a}^{-1}$ \\
$\Delta x$  & representative width (diameter) of map-plane mesh cells & km
\end{tabular}
\end{center}

\medskip
\caption{Notation for performance modeling.  Parameters $\delta,\eps,m, m_c,n$ are pure numbers.}
\label{tab:notation}
\end{table*}

@@

\newcommand{\oo}[1]{{\LARGE \strut} \displaystyle O\left(#1\right)}
\setlength{\tabcolsep}{5pt}
\renewcommand{\arraystretch}{1.5}
\begin{table*}[ht]
\begin{center}
\begin{tabular}{lllll}
\emph{time-stepping} & \emph{dynamics} & \emph{work in 1D} & \emph{work in 2D} & \emph{caveat} \\ \hline
explicit & SIA    & $\oo{\frac{D}{L^2} n^3}$       & $\oo{\frac{D}{L^2} n^2}$ \\
explicit & Stokes & $\oo{\frac{D}{L^2} n^{3+\delta}}$ & $\oo{\frac{D}{L^2} n^{2+\delta}}$ &  $D$ effects dominant \\
explicit & Stokes & $\oo{\frac{U}{L} n^{2+\delta}}$    & $\oo{\frac{U}{L} n^{1.5+\delta}}$  & $U$ effects dominant \\
implicit & SIA    & \multicolumn{2}{c}{$\oo{q\, n^{1+\eps}}$} \\
implicit & Stokes & \multicolumn{2}{c}{$\oo{q\, n^{1+\eps+\delta}}$}
\end{tabular}
\end{center}

\medskip
\caption{Asymptotic scaling of computational work, in flops per model year, for time-stepping numerical ice sheet simulations over domain width $L$, as the degrees of freedom $n$ go to infinity (high horizontal resolution).  See Table \ref{tab:notation} for remaining notation; \emph{1D} and \emph{2D} refer to flow-line and map-plane simulations, respectively.}
\label{tab:performancemodel}
\end{table*}

>>>>

\title[Performance models for high-resolution ice sheet simulations]{Letter: Performance models for \\ high-resolution ice sheet simulations}

@@

\begin{table*}[ht]
\begin{tabular}{cll}
\emph{name} & \emph{meaning} & \emph{units} \\ \hline
$\alpha$    & one fixed-geometry Stokes velocity solution requires $O(n^{1+\alpha})${\large \strut} work\\
$\beta$     & one implicit SIA geometry-update requires $O(n^{1+\beta})$ work \\
$\gamma$    & one implicit Stokes geometry-update plus velocity solution requires $O(n^{1+\gamma})$ work \\
$D$         & representative geometric (shallow ice approximation) diffusivity of an ice sheet & $\text{km}^2 \text{a}^{-1}$ \\
$L$         & width of simulation domain & km \\
$n$         & degrees of freedom: number of nodes in the horizontal (flowline or map-plane) mesh \\
$q$         & time steps per model year needed to resolve modeled climate interactions & $\text{a}^{-1}$ \\
$T$         & simulation duration in model years & a \\
$\Delta t$  & length of time step in model years & a \\
$U$         & representative horizontal ice velocity & $\text{km}\,\text{a}^{-1}$ \\
$\Delta x$  & representative width (diameter) of map-plane mesh cells & km
\end{tabular}
\caption{Notation for performance modeling.  Parameters $\alpha,\beta,\gamma,n$ are pure numbers.}
\label{tab:notation}
\end{table*}

@@

\newcommand{\oo}[1]{{\LARGE \strut} \displaystyle O\left(#1\right)}
\setlength{\tabcolsep}{5pt}
\renewcommand{\arraystretch}{1.5}
\begin{table}[ht]
\begin{tabular}{lll}
\emph{time-stepping} & \emph{dynamics} & \emph{work (flops per model year)} \\ \hline
explicit & SIA    & 1D:\quad $\oo{\frac{D}{L^2} n^3}$ \\
         &        & 2D:\quad $\oo{\frac{D}{L^2} n^2}$ \\
explicit & Stokes & 1D:\quad $\oo{\frac{D}{L^2} n^{3+\alpha} + \frac{U}{L} n^{2+\alpha}}$ \\
         &        & 2D:\quad $\oo{\frac{D}{L^2} n^{2+\alpha} + \frac{U}{L} n^{1.5+\alpha}}$ \\
implicit & SIA    & $\oo{q\, n^{1+\beta}}$ \\
implicit & Stokes & $\oo{q\, n^{1+\gamma}}$
\end{tabular}
\caption{Asymptotic scaling of computational work for time-stepping numerical ice sheet simulations as the degrees of freedom $n$ go to infinity (high horizontal resolution).  Note 1D and 2D refer to flow-line and map-plane simulations, respectively.  See Table \ref{tab:notation} for remaining notation.}
\label{tab:performancemodel}
\end{table}
